---
title: "Project2"
author: "Serg C"
date: "August 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rep Course Project 2

This project includes several parts. 

## 1. Reading the dataset

The dataset is downloaded and unzipped, and then `dt` data.frame is read from `activity.csv` file:

```{r load.1, echo=FALSE}
wdir <- '.'
fZip <- paste(wdir,'activity.zip',sep="/")
if(!file.exists("activity.csv")){
 if(!file.exists(fZip)){
  fUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  download.file( fUrl, fZip )
 }
 if(!file.exists(paste(wdir,"activity.csv",sep="/"))){
  unzip( fZip, exdir=wdir )
 }
}
```
```{r read}
dt <- read.csv( paste( wdir,"activity.csv",sep="/") )
```

## 2. Histogram of the total number of steps taken each day

Calculate the total number of steps taken per day, ignore NAs.

```{r calc.2}
tot <- tapply( dt$steps,dt$date, sum, na.rm=TRUE )
```

Make a histogram of the total number of steps taken each day

```{r hist.2}
hist( tot, main='Histogram of Total Steps per Day',xlab='Total Steps')
```

## 3. Mean and median number of steps taken each day

Calculate and report the mean and median of the total number of steps taken per day.

The report record is formatted as < date, median, mean, max >

```{r calc.3}
rep <- aggregate( steps~date, data=dt, function(x){summary(x, digits=2)})
day <- data.frame( unclass(rep), check.names = FALSE, stringsAsFactors = FALSE)[,c(1,4,5,7)]
names(day) <- c('date','median','mean','max')
print(day,row.names=F)
```

## 4. Time series plot of the average number of steps taken

Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

Calculate
```{r calc.4}
rep <- aggregate( steps~interval, data=dt, mean, na.rm=TRUE )
```

then plot
```{r plot.4}
with(rep,plot(interval,steps,type='l',
    main='Daily Average Steps per Interval',ylab='Average Steps',xlab='Interval'))
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r which.5}
int <- rep[which(rep$steps==max(rep$steps)),"interval"]
```

**Q:** Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

**A:** The interval that contains the maximum number of steps on average is ***`r int`***.
```{r print.5, echo=FALSE}
print(rep[which(rep$steps==max(rep$steps)),],row.names=F)
```

## 6. Code to describe and show a strategy for inputing missing data

1. Calculate and report the total number of missing values in the dataset

```{r calc.6}
dt_na <- is.na(dt$steps)
print( sum(dt_na), row.names=F )
```

2. My strategy to fill in NA's is replace NA Steps observation with the average of value of Steps claculated for the same interval in item 4.  Wewill merge the original `dt` and `rep` frames, then re-set NA's then make a new dataset.  

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r create.6}
dt.new <- merge( dt, rep, by='interval' )
dt.new[ is.na(dt.new[,2]), 2] <- dt.new[ is.na(dt.new[,2]), 4]
dt.new <- dt.new[,c(2,3,1)]
names(dt.new)<-c('steps','date','interval')
```

## 7. Histogram of the total number of steps taken each day after missing values are fixed

Calculate the total number of steps taken per day.

```{r calc.71}
tot <- tapply( dt.new$steps, dt.new$date, sum )
```

Make a histogram of the total number of steps taken each day (fixed data frame)

```{r hist.72}
hist( tot, main='Histogram of Total Steps per Day',xlab='Total Steps')
```

Calculate and report the mean and median of the total number of steps taken per day.

The report record is formatted as < date, median, mean, max >

```{r calc.73}
rep <- aggregate( steps~date, data=dt.new, function(x){summary(x, digits=2)})
day <- data.frame( unclass(rep), check.names = FALSE, stringsAsFactors = FALSE)[,c(1,4,5,7)]
names(day) <- c('date','median','mean','max')
print(day,row.names=F)
```

**Q:** Do these values differ from the estimates from the first part of the assignment? 

**A:** Yes


**Q:** What is the impact of fixing missing data on the estimates of the total daily number of steps?

**A:** The total daily number of steps has increased.

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Modify`dt.new` data frame to add a new column, calculate df to plot

```{r calc.8}
dt.new$day <- ifelse(
    weekdays(as.Date(dt.new$date,'%Y-%m-%d')) %in% c('Sunday','Saturday' ),
    'weekend', 'weekday')
wd <- aggregate( steps~interval, data=subset(dt.new, day=='weekday'), mean )
we <- aggregate( steps~interval, data=subset(dt.new, day=='weekend'), mean )
```

Make a panel plot

```{r plot.8}
par(mfrow = c(2, 1))
with(we,plot(interval,steps,type='l',
    main='Weekend Average Steps per Interval',ylab='Average Steps',xlab='Interval'))

with(wd,plot(interval,steps,type='l',
    main='Weekday Average Steps per Interval',ylab='Average Steps',xlab='Interval'))
```

## 9. URL to GitHub report

[rep_proj_2.html] (http://htmlpreview.github.com/?https://github.com/about-r/RepData_PeerAssessment1/master/rep_proj_2.html)
